

##  Movie ratings and lengths (13.1)


 Before a movie is shown in theatres, it
receives a "rating" that says what kind of material it
contains. [link](https://en.wikipedia.org/wiki/Motion_Picture_Association_of_America_film_rating_system)
explains the categories, from G (suitable for children) to R (anyone
under 17 must be accompanied by parent/guardian). In 2011, two
students collected data on the length (in minutes) and the rating
category, for 15 movies of each rating category, randomly chosen from
all the movies released that year. The data are at
[link](http://ritsokiguess.site/datafiles/movie-lengths.csv).



(a) Read the data into R, and display (some of) what you read in.

```{r}
library(tidyverse)
my_url <- "http://ritsokiguess.site/datafiles/movie-lengths.csv"
movies <- read_csv(my_url)
movies
```


(b) Count how many movies there are of each rating.







(d) Make a graph to assess whether this ANOVA is
trustworthy. Discuss your graph and its implications briefly.


```{r}
ggplot(movies, aes(x = rating, y = length)) + geom_boxplot()
```

look at the bootstrap sampling distribution of the sample mean for any groups that concern us.

Work one group at a time (eg PG):

```{r}
movies %>% filter(rating == "PG") -> pgs
pgs
```

```{r}
tibble(sim = 1:1000) %>% 
  rowwise() %>% 
  mutate(my_sample = list(sample(pgs$length, replace = TRUE))) %>% 
  mutate(my_mean = mean(my_sample)) %>% 
  ggplot(aes(sample = my_mean)) + stat_qq() + stat_qq_line()
```

The G group goes similarly:

```{r}
movies %>% filter(rating == "G") -> gs
gs
```

and

```{r}
tibble(sim = 1:10000) %>% 
  rowwise() %>% 
  mutate(my_sample = list(sample(gs$length, replace = TRUE))) %>% 
  mutate(my_mean = mean(my_sample)) %>% 
  ggplot(aes(sample = my_mean)) + stat_qq() + stat_qq_line()
```

this one is skewed to the left (bunched up at the top, though not very spread out at the bottom, surprisingly)