## Number 1 songs (problem 17.7)

```{r}
library(tidyverse)
```


 The data file
[link](http://stat405.had.co.nz/data/billboard.csv) contains a lot of
information about songs popular in 2000. This dataset is untidy.  Our
ultimate aim is to answer "which song occupied the \#1 position for the largest number of weeks?". To do that, we will build a pipe that
starts from the data frame read in from the URL above, and finishes
with an answer to the question. I will take you through this step by
step. Each part will involve adding something to the pipe you built
previously (possibly after removing a line or two that you used to
display the previous result).



(a) Read the data and display what you have.

```{r}
my_url <- "http://stat405.had.co.nz/data/billboard.csv"
songs <- read_csv(my_url)
songs
```

There is one row per song, so 317 *songs* total. (the "week" columns with the rank on the chart for that week were probably called "1st week", "2nd week" etc in the original spreadsheet; column names in R cannot start with a number, so `x` got added to the front when they were read in. If you open the URL in your browser, it will probably open up a spreadsheet app, and you can see whether I was right.)


(b) The columns `x1st.week` through
`x76th.week` contain the rank of each song in the Billboard
chart in that week, with week 1 being the first week that the song
appeared in the chart.  Convert all these columns into two: an
indication of week, called `week`, and of rank, called
`rank`. Most songs appeared in the Billboard chart for a
lot less than 76 weeks, so there are missing values, which you
want to remove.  (I say "indication of week" since this will
probably be text at the moment). Display your new data frame. Do
you have fewer columns?  Why do you have a lot more rows? Explain
briefly.

```{r}
songs %>% pivot_longer(starts_with("x"), names_to = "week", values_to = "rank",
                       values_drop_na = TRUE)
```

There are now only 9 columns, way down from the original 83, because all the 76 week columns have been condensed into 2. In this dataframe, there are many more rows because now each "observation" is a particular song in a particular week, and these songs appeared in the Billboard chart for several weeks each.

(c) Both your `week` and `rank` columns are
(probably) text. Create new columns that contain just the numeric
values, and display just your new columns, again adding onto the
end of your pipe. If it so happens that `rank` is already a number, leave it as it is.

Here, the `rank` came out as text, so we don't need to worry about that.

Get the number out of `week`, and save it back into `week` (or create a new column if that is scary:

```{r}
songs %>% pivot_longer(starts_with("x"), names_to = "week", values_to = "rank",
                       values_drop_na = TRUE) %>% 
  mutate(week = parse_number(week))
```

The `week` is now a number.

(e) The meaning of your week-number column is that it refers
to the number of weeks *after* the song first appeared in the
Billboard chart. That is, if a song's first appearance (in
`date.entered)` is July 24, then week 1 is July 24, week 2
is July 31, week 3 is August 7, and so on. Create a column
`current` by adding the appropriate number of *days*,
based on your week number, to `date.entered`. Display
`date.entered`, your week number, and `current` to
show that you have calculated the right thing. Note that you can
add a number of days onto a date and you will get another date.

`current` should contain entered date plus the number of weeks *minus one*, times 7 to make days.

```{r}
songs %>% pivot_longer(starts_with("x"), names_to = "week", values_to = "rank",
                       values_drop_na = TRUE) %>% 
  mutate(week = parse_number(week)) %>% 
  mutate(current = date.entered + (week - 1) * 7) -> songs.2
songs.2
```

I decided that this is the end of the tidying, so I gave this dataframe a name.

(f) Reaching the \#1 rank on the Billboard chart is one of
the highest accolades in the popular music world. List all the
songs that reached `rank` 1. For these songs, list the
artist (as given in the data set), the song title, and the date(s)
for which the song was ranked number 1. Arrange the songs in date
order of being ranked \#1. Display all the songs (I found 55 of them).

This is just a `filter`, followed by a `select` to get only the columns we want:

```{r}
songs.2 %>% filter(rank == 1) %>% 
  select(artist.inverted, track, current) %>% 
  arrange(current) -> songs.3 
songs.3
```

55 indeed.

(g) Use R to find out which song held the \#1 rank for the
largest number of weeks. For this, you can assume that the song
titles are all unique (if it's the same song title, it's the same
song), but the artists might not be (for example, Madonna might
have had two different songs reach the \#1 rank). The information
you need is in the output you obtained for the previous part, so
it's a matter of adding some code to the end of that.
The last mark was for displaying *only* the song that was
ranked \#1 for the largest number of weeks, or for otherwise
making it easy to see which song it was.

Use the dataframe we just made, count the number of rows that each song appears on, then sort the counts in descending order:

```{r}
songs.3 %>% 
  count(track) %>% 
  arrange(desc(n))
```

This loses the artist name. If you want to keep it, make a combined song + artist name first, and then count and sort:

```{r}
songs.3 %>% 
  mutate(song = str_c(track, " by ", artist.inverted)) %>% 
  select(song, current) %>% 
  count(song) %>% 
  arrange(desc(n))
```

This protects you against two different songs with the *same* title both making it to number 1; the artist would probably be
different (eg., a cover version of a song).

