---
title: "Soap"
output: html_notebook
---

## packages

```{r}
library(tidyverse)
```

## the soap data

We read in the data

```{r}
my_url <- url("http://ritsokiguess.site/datafiles/soap.txt")
soap <- read_delim(my_url, " ")
soap
```


There are 27 observations on 4 variables.

## Some graphs

A histogram of the `scrap` values

```{r}
ggplot(soap, aes(x = scrap)) + geom_histogram(bins = 10)
```

Boxplot of scrap by line

```{r}
ggplot(soap, 
       aes(x = line, 
           y = scrap)) + 
  geom_boxplot()
```

or summarize  by numbers (from lecture 2b)

```{r}
soap %>% 
  group_by(line) %>% 
  summarize(med_scrap = median(scrap), 
            med_speed = median(speed))
```


Scrap and speed: scatterplot

```{r}
ggplot(soap, aes(x = speed, y = scrap)) + geom_point()
```

If the production line is run faster, there tends to be more scrap.

Can we indicate which production line each observation came from? Use colour:

```{r}
ggplot(soap, aes(x = speed, y = scrap, colour = line)) + geom_point()
```

the red points are usually above the blue ones, so line A produces more scrap on average than line B at the same speed.

```{r}
ggplot(soap, aes(x = speed, y = scrap, colour = line)) + geom_point() + geom_smooth(method = "lm")
```






